The Resurgence of the Linear Optics Interferometer --- Recent Advances \& Applications

### 2017-06-29 Review merged 30-4-2017/paper.tex

\comment{To do}

@@

\section{Mathematical background}

A idealized single photon in a quantum interferometer is described by its creation operator $\adagh_{j}$, where $j$ is the label of the mode the photon is in  within the interferometer. The creation and annihilation operators satisfy the bosonic comutation relationship $[\ah_{j},\adagh_{k}]=\delta_{j,k}$. A similar commutation relationship can be written when additional degrees of freedom, such as polarization, orbital angular momentum, and time-bins \cite{bib:Tillmann2015,bib:Bozinovic2013, bib:Nicolas2014, bib:Humphreys2013, bib:Donohue2013}, are present. When multiple photons are present, they experience quantum interference when all quantum labels are the same.

@@

\begin{align}
b_i^{\dag}=\sum_{j=1}^d U_{ij}a_j^\dag \ ,
\end{align}
where $a_j^\dag$ and $b_i^\dag$ are the creation operators of a single input and output photon in the $j$-th and $i$-th modes respectively, and $U\in SU(d)$. All such transformations can be expressed as sequences of beamsplitters and phase-shifters \cite{bib:Reck1994}. In the case when photons have additional labels, for instance, if they have internal labels on top of spatial labels, it is also possible to derive an analogous decomposition, known as a cosine-sine decomposition \cite{bib:Dhand2015}, that realizes the unitary transformation on the photons as a sequence of beamsplitters and internal transformations. Toolkits using group theory are being developed to deal with partial distinguishabilities among interfering photons \cite{bib:Tan2013,bib:deGuise2014,bib:deGuise2015}. Others use quantum-to-classical transitions to explain multiparticle interference \cite{bib:Ra2013}. \sihui{Experimental implementation [Walmsley, Jennewein]}

@@

Time-bin encoding was first conceived via a single-photon passing through a Mach-Zedner interferometer with its two paths having different lengths \cite{bib:Brendel99}. If the photon passes through the shorter (longer) path, it will arrive at the output port in an `early' (`later') time-bin labelled $\ket{e}$ ($\ket{\ell}$). Thus, the photon will exit the interferometer in a state that is a superposition of these two states. Owing to difficulties in implementing qubit operations in this basis, it was at the time mostly used for demonstrating quantum communication over long distances \cite{bib:Thew02,bib:Marcikic04}. With the advent of faster optical components, and single-photon detectors, it has become feasible to perform any single qubit operation, and a post-selected CPHASE gate on time-bin qubits \cite{bib:Humphreys2013}. These gates form an universal set for quantum computing. At the same time, an ultrafast measurement technique for recovering time-bin qubits was also demonstrated \cite{bib:Donohue2013}.

In Humphrey {\it et al.} \cite{bib:Humphreys2013}, the time-bin qubits have a polarization register which is used to toggle between the mode in which qubits are stored and transmitted, {\it i.e.} the `register' mode, and the mode in which qubits are manipulated, {\it i.e.} the `processing' mode using a polarization rotation. Other operations needed are a displacement operation, that moves a time-bin qubit in the processing polarization in time relative to other qubits, a partial polarization rotation between the register and processing polarizations to couple them, and a photon number measurement in a bin. The polarization rotation and polarization coupling operations were implemented using a fast integrated optical switch based on polarization-sensitive cross-phase modulation with a switching time of less than 10ps. The displacement was implemented using a calcite crystal of a size that is of a path-length difference equal to an integer multiple of the time-bin separation. Using these operations and post-selection using a single-photon detectors, the authors were able to perform a heralded CPHASE gate \cite{bib:KLM01}.

In Donohue {\it et al.} \cite{bib:Donohue2013}, time-bins were converted into frequency bins using a nonlinear optical process known as sum-frequency generation (SFG) by pumping them on a crystal with a laser beam. The waveforms of the input photons and the pump beam are chirped in opposite directions in frequency, so that the output is a set of two peaks separated in frequency by an amount proportional to the time separation of the time bins. The output can then be measured with detectors that are slow compared to this separation. This is unlike conventional time-bin measurements which involve sending a time-bin state through an unbalanced Mach-Zedner interferometer matched to the time-bin separation to produce three output pulses that are separated by the time delay, and then having to resolve the middle pulse in time.

Apart from these promising advances in implementing time-bin qubit gates and ultrafast time-bin measurements, scalable networks using time-bin encoding are also possible using a loop-based architecture \cite{bib:Motes14}. This is discussed in more detail in Section \ref{sec:timebins}.

@@

The task of implementing an arbitrary quantum computation on linear optics comes down to implementing an arbitrary $n\times n$ unitary matrix. If a non-unitary transformation is desired, it can be embedded within a unitary matrix with larger dimensions. An algorithm for expressing an arbitrary unitary matrix {\it exactly} in terms of a sequence of beamsplitters and phase-shifters was described by Reck \emph{et al.} \cite{bib:Reck1994}. This decomposition requires $\mathcal{O}(n^2)$ linear optical elements, and the algorithm for finding the decomposition has polynomial runtime. Thus, such decompositions can always be determined and implemented efficiently. The layout for the original Reck \emph{et al.} decomposition is shown in Fig.~\ref{fig:Reck}. However, since then a multitude of alternate decompositions have been found. A notable downside of the original decomposition is that different photons experience different circuit depth, i.e pass through different numbers of optical elements, resulting in asymmetry in losses and the accumulation of errors.

\begin{figure}[!htb]
\includegraphics[width=\columnwidth]{reck}
\caption{Efficient decomposition of arbitrary linear optics networks into a sequence of beamsplitters and phase-shifters.} \label{fig:Reck}	
\end{figure}

Alternatively, Mach-Zedner interferometers can also be employed as building blocks instead of beamsplitters and phase-shifters \cite{bib:Reck1994, bib:Englert2001}. Later, it has been shown that any nontrivial beamsplitter, that does more than permuting modes or adding phases to them, is universal for linear optics \cite{bib:Bouland2014}. However, they do not provide an explicit construction for arbitrary unitaries.

If the linear optical transformations can be realized on various degrees of freedom of light, then it is possible to realize a $n\times n$ arbitrary unitary transformation, where $n=n_s n_p$ for $n_s$ spatial modes, and $n_p$ internal modes, by a sequence of $\mathcal{O}(n_s^2 n_p)$ beamsplitters and $\mathcal{O}(n_s^2)$ internal transformations \cite{bib:Dhand2015}. Their approach reduces the required number of beamsplitters but increases the total number of optical elements needed by a factor of 2.

@@

In many practical situations, the structure of a linear optical device in terms of its constituent beamsplitters and phase-shifters is known once it is built. However, owing to manufacturing imperfections, a precise characterization of these devices is still needed post-production. One approach for achieving this is via quantum process tomography \cite{bib:Mitchell03,bib:Obrien04,bib:Lobino08,bib:Saleh11}. However, quantum process tomography is an expensive approach in terms of the number of measurements required to characterize the network, with exponential overhead, becoming impractical for large optical networks which can be as large as $n=900$ modes using present-day technology \comment{(check citations for number of modes)}. However, alternative characterization protocols have been developed using quantum interference of various quantum light sources \cite{bib:Laing12,bib:Rahimi-Keshari13} in the linear optical device.

Generally, the unitary matrices of $d\times d$ linear optical devices are complex $U_{ij}=r_{ij}e^{i\theta_{ij}}$, where $0 \leq r_{ij}\leq 1$, and $0\leq \theta_{ij}\leq 2\pi$. The scheme in \cite{bib:Laing12} relies on injecting one- and two-photon states into the network with correlated photon detection. First, they note some equivalencies: two unitaries $U$ and $U'$ are equivalent if there exist two diagonal unitary matrices $D_1^U$ and $D_2^U$ such that $U'=D^U_1 U D^U_2$, because these diagonal matrices are regarded as unknown and trivial phases on the input and output ports of the network, to which the one- and two-photon data are insensitive to. This reduces the first row and column elements to real numbers, i.e. $\theta_{1j}=\theta_{i1}=0$. Second, the photo-statistics remain invariant under complex conjugation by $U$. Thus, the imaginary part of $M_{2,2}$ must be non-negative. Then, assuming the first two rows and columns are non-vanishing, and there is no total loss in the interferometer, the matrix to be recovered is
\begin{align}
U=\left (\begin{array}{cccc}
r_{11} & r_{12} & \ldots & r_{1m} \\
r_{21} & r_{22}e^{i\theta_{22}} &\ldots & r_{2m}e^{i\theta_{2m}}\\
\vdots & \vdots & \ddots & \vdots \\
r_{m1} & r_{m2}e^{i\theta_{m2}} & \ldots & r_{mm}e^{i\theta_{mm}}
 \end{array}\right ) \ .
\end{align}
They showed it is possible to write the parameters $r_{ij}$, and $\theta_{ij}$ for $i,j\geq 2$ in terms of the $2m-1$ real parameters of the first column and row, and the visibility of two-photon inputs. The remaining $2m-1$ real parameters can be found via one-photon 
transmissions. An increased accuracy in the characterization is possible by estimating and correcting systematic errors that arise due to mode mismatch \cite{Dhand16}. Others have used numerical methods to find the closest parameters that yield the observed visibilities \cite{bib:Spagnolo16,bib:Tillmann16}.

Another characterization method was presented that is similar to \cite{bib:Laing12} with the important exception that coherent states are to be used instead of Fock states \cite{bib:Rahimi-Keshari13,bib:Heilmann15}. Such states are produced by a standard laser source, thus reducing experimental resources. The $r_{jk}$ terms are found by the square root of the ratios of output intensities at the $k$th port to the input intensity at the $j$th port. The remaining phases $\theta_{ij}$ are found by the interference pattern given by a two-mode coherent state $\ket{\alpha_1}\ket{\alpha_2}$ created by splitting a single coherent state on a 50-50 beamsplitter, and then imparting a relative phase, $\phi$, between them. The states $\ket{\alpha_1}$ and $\ket{\alpha_2}$ are input into port 1, and $j$ respectively. The output intensity at the $j$th port is then
\begin{align}
I_k=I(r_{1k}^2+r_{jk}^2+2 r_{1k}r_{kj}\cos(\phi+\theta_{jk})) \ ,
\end{align}
where $\theta_{jk}=0$ for $k=1$. By scanning the phase shift $\phi$ and locating the maximum value of $I_k$ for $j=2,\ldots, m$, all unknown phases can be found via $\theta_{jk}=2\pi-\phi$. An elegant extension of the scheme of Rahimi-Keshari {\it et al.} removes the need for precise control of the phase-shift $\phi$ \cite{bib:Heilmann15} by suggesting instead to plot the output intensity $I_k$ with respect to the input intensity $I$. In time, the natural drift in the laser source will cause this plot to trace out an ellipse, known as a Lissajous figure, whose orientation and direction of evolution will give the phase $\theta_{jk}$ and its sign respectively.

@@

Significant progress has been made in the experimental realisation of linear optics protocols, and have become standard and widespread. We now discuss some of these advances.

@@

The photonic states that are most commonly employed in linear optics protocols can be divided into Fock states (e.g single-photon), and entangled states, such as EPR, GHZ and cluster states. We will consider advances in each of these.

@@

Sources of single photons for applications in quantum information processing can be separated into two main categories \cite{bib:Kok05}; those produced by spontaneous parametric down-conversion (SPDC), and those by solid-state emitters in a cavity. Both categories have seen major development recently in photon-number, with high fidelities. To-date, SPDC has been able to produce up to ten entangled photons \cite{bib:WangChen16,bib:Chen17}, and five-photon quantum interference has been reported using a quantum dot emitter in a microcavity \cite{bib:WangHe16}.

In SPDC, a nonlinear crystal with a large $\chi^{(2)}$ non-linearity is pumped with a strong coherent state (i.e laser source) and with a small probability, the pump beam is absorbed by the crystal to produce two beams of lower energy known as the signal and idler. Owing to conservation of energy and momentum, the two beams have spatio-temporal correlations that can be engineered to produce twin-beam states with perfect  photon number correlation, of the form
\begin{align}
\ket\psi_\mathrm{SPDC} = \sqrt{1-\chi^2} \sum_{n=0}^\infty \chi^n \ket{n,n},	
\end{align}
where $\chi$ is the squeezing parameter. If a single photon were to be detected in one of the modes, it is certain that the other mode would similarly contain a single photon.

Some commonly used nonlinear crystals for SPDC are beta barium borate (BBO), periodically poled lithium niobate (PPLN), and periodically poled  potassium titanyl phosphate (PPKTP). Recently, techniques using bismuth triborate (BiBO) have improved to the extent of becoming one of the record-holders in single-photon production \cite{bib:WangChen16}.

Solid-state single photon sources come from semiconducting nanostructures and nitrogen valencies (NV) centers in diamond. Both types are versatile and efficient sources of single photons, however, the photons they produce have suffered from the lack of indistinguishability that is necessary for typical quantum information processing applications. Recent developments in the former have been very promising. Notably, much progress has been made using resonant excitation of quantum dots. Here, laser pulses are used to excite their electronic resonance and trigger the emission of single photons \cite{bib:Muller07,bib:Vamivakas09, bib:Flagg09,bib:Ates09, bib:Dirk10,bib:1748,bib:Jayakumar13,bib:Wei14,bib:Muller14,bib:Unsleber15,
bib:237403,bib:Sweeney14}. An architecture that embeds the quantum dot in a micropillar cavity with the same resonant frequency as the dot exploits the Purcell effect to achieve higher single-photon production efficiency€ \cite{bib:Ding16,bib:213601}. Such resonant excitation of quantum dots overcomes the homogenous broadening of the excited state that causes degradation of photon purity and improving indistinguishability. Using a time-bin loop-based architecture \cite{bib:Motes14} on a stream of single photons produced by such a quantum dot, the same group was able to demonstrate quantum interference between five indistinguishable photons \cite{bib:WangChen16}.

@@

An EPR pair, or Bell pair, is one of the four states,
\begin{align}
\ket{0}_A\ket{0}_B &\pm \ket{1}_A\ket{1}_B, \nonumber \\
\ket{1}_A\ket{0}_B &\pm \ket{0}_A\ket{1}_B,
\end{align}
which are all maximally entangled and locally equivalent to one another. These are the simplest examples of entangled states, and their preparation via SPDC has been the mainstay of entangled state preparation for quantum optical processing \cite{bib:Kim01,bib:Kim03,bib:Brida07,bib:Dayan07}. In some applications, it may be desired to have the EPR pairs conditionally prepared, {\it i.e.} successfully prepared only under certain measurement outcomes of auxiliary modes. Several theoretical approaches have been proposed for this purpose \cite{bib:Pittman03,bib:Sliwa03,bib:Walther07}, and demonstrated \cite{bib:Wagenknecht10,bib:Barz10}. They are often used as a building block for preparing other multipartite entangled states.

@@

GHZ states form a class of entangled quantum states on multiple subsystems with at least three parties \cite{bib:GHZ89}. For qubit encodings with $n$ subsystems, a GHZ state is of the form
\begin{align}
\ket{\Psi_{n}^{\rm (GHZ)}}=\frac{\ket{0}^{\otimes n}+\ket{1}^{\otimes n}}{\sqrt{2}} \ .
\end{align}
These states are also non-local, and have been used extensively in experimental test for non-locality \cite{bib:JW00,bib:Zhang15}. This has been a subject matter covered in detail by another review \cite{bib:JW12}.

@@

Cluster states form another class of multiparty entangled states, and were conceived in the context of arrays of qubits with an Ising-type interaction \cite{bib:Briegel01}.

These states can be represented as a graph, in which vertices are qubits initialized into the superposition state $\ket{+}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{1})$, with CPHASE operation applied between edges, as shown in Fig.~\ref{fig:cluster}. For this reason, such states are also referred to as graph states. The CPHASE operations generate entanglement between the qubits, and because they commute, are independent of ordering. Cluster states are a resource state for a model of quantum computation known as measurement-based quantum computation (MBQC). Here, having such a state as a resource enables universal quantum computation using only single-qubit measurements \cite{bib:Raussendorf03}. Therefore, the preparation of such states is highly valuable, generating much interest in their efficient preparation.

\begin{figure}[!htb]
\includegraphics[width=0.6\columnwidth]{cluster_state}
\caption{The representation of cluster states as a graph. Vertices represent qubits initialised into the $\ket{+}$ state, while edges represent the application on CPHASE gates, the ordering of which is irrelevant.} \label{fig:cluster}	
\end{figure}

Unfortunately, implementing CPHASE gates using linear optics is complicated and highly non-deterministic. A major improvement upon this is to use fusion gates - rotated polarising beamsplitters, which implement projections onto the Bell basis. These operations fuse smaller cluster states, represented using polarisation encoding, into larger ones, consuming one (type-I fusion) or two (type-II fusion) photons in the process. These operations are shown in Fig.~\ref{fig:fusion}. Although these operations are non-deterministic with a success probability of $1/2$, they require only a single beamsplitter, already a major improvement over directly implementing CPHASE gates. Furthermore, unlike CPHASE gates, fusion operations require only high Hong-Ou-Mandel visibility, rather than the Mach-Zehnder stability required within existing CPHASE gate implementations, a major experimental simplification.

Although these gates are non-deterministic and consume qubits, strategies have been described for efficiently preparing arbitrarily large cluster states of arbitrary topology, and using far fewer optical elements than by directly employing CPHASE gates.

\begin{figure}[!htb]
\includegraphics[width=0.7\columnwidth]{fusion}
\caption{Fusion gates for joining smaller cluster states into larger ones. Both require only a single polarising beamsplitter, and waveplates. The type-I and -II fusion gates consume 1 or 2 qubits (photons) respectively, creating an edge between the remaining graphs. The type-I gate consumes one fewer photon, but requires number-resolved detection on the detected mode. The type-II gate requires only on/off photodetection, but consumes an additional photon. Thus, the type-I gate can be employed to fuse two Bell pairs into a 3-qubit cluster state, whereas the type-II gate will only grow clusters when beginning with at least 3 qubits per cluster.} \label{fig:fusion}	
\end{figure}

@@

Numerous experimental implementations of various linear optics protocols have been demonstrated, using a variety of architectures. The main contenders considered to date are: bulk-optics; waveguides; and time-bin encoding.

@@

The early implementations of linear optics protocols relied on bulk-optics implementations, in which a circuit is decomposed into a discrete arrangement of optical components, specifically beamsplitters and phase-shifters. This corresponds to a direct implementation of the Reck \emph{et al.} decomposition shown in Fig.~\ref{fig:Reck}, or subsequent alternate decompositions.

Although this approach is effective and conceptually straightforward, it is highly impractical for large interferometers. For example, the implementation of a 100-mode interferometer requires on the order of 10,000 discrete optical elements, which must all be simultaneously aligned with interferometric stability, a highly experimentally challenging problem, not to mention incredibly physically large.

@@

Alternately, integrated waveguide devices can be employed to implement identical decompositions by etching the modes into paths in an optical chip, where evanescent coupling between closely neighbouring modes implements beamsplitter operations. Such devices are extremely compact, allowing the miniaturisation of large numbers of optical elements. Importantly, such devices are extremely optically stable, mitigating the need for any kind of dynamic stabilisation

@@

@@

The workings of the scheme can be thought of as follows: the inner loop allows arbitrary beamsplitter operations between neighbouring time-bins; the outer loop does nothing interferometric, but rather enables the pulse-train to undergo as many applications of the inner loops as necessary. It then follows that this scheme is universal for linear optics, as a sufficient number of beamsplitter operations between neighbouring modes enables universal decompositions, using for example, the Reck \emph{et al.} decomposition described earlier.

As a simple example of how such time-bin encoding maps to other encodings, in Fig.~\ref{fig:PBS_TB} we show the isomorphism between the polarising beamsplitter operation and a pairwise temporal beamsplitter operation. This implies a direct mapping for implementing cluster state preparation within the time-bin scheme.

@@

\subsection{Photodetection}

Broadly speaking, there are two main classes of photodetectors: number resolved, and non-number-resolved (also referred to as bucket, or on/off) detectors. As the names suggest, these detectors are able to measure photon-number, or merely the presence or absence of photons. They are described by the measurement projectors,
\begin{align}
\hat\Pi_n &= \ket{n}\bra{n}, \nonumber \\
\hat\Pi_\text{off} &= \hat\Pi_0, \nonumber \\
\hat\Pi_\text{on} &= \hat{I} - \hat\Pi_\text{on},	
\end{align}
respectively.

The dominant challenge facing the experimental implementation of both types of detectors is inefficiency, which results in higher photon-number being confused as having lower photon-number. In some protocols, such as type-II fusion, which relies on coincidence detection, inefficiency is not problematic. However, in most applications, including type-I fusion, this is highly problematic as photon-number confusion results in decoherence via projecting onto the wrong photon-number, unknown to the observer.

\comment{What are the different technologies}

@@

@@

Linear optics interferometry has become a leading contender for the implementation of quantum information processing protocols. The preparation, manipulation and measurement of photonic quantum information has become mainstream and widely employed, with impressive experimental accuracies.

We have discussed the various ways in which quantum information can be represented photonically, how they can be manipulated and detected, and some of their leading applications.

Although there are countless physical architectures for the implementation of quantum information processing, and it is far from certain which will win `the quantum race for supremacy', optics will always find a home as the only contender for applications involving quantum communication. For this reason, the future of linear optics interferometry is a bright one, which will find inevitable applicability in the future quantum world.

>>>>

\sihui{Si-Hui can colour code things she adds like this}

\peter{And Peter can do it like this}

\comment{Let's add comments and questions like this}

@@

\section{Mathematical background}
\comment{Mathematical representation for LO networks, and very basic background on quantum optics}\\

A idealized single photon in a quantum interferometer is described by its creation operator $\adagh_{j}$, where $j$ is the label of the mode the photon is in  within the interferometer. The creation and annihilation operators satisfy the bosonic comutator relationship $[\ah_{j},\adagh_{k}]=\delta_{j,k}$. A similar commutator relationship can be  written up when more degrees of freedom, such as polarization, orbital angular momentum, and time-bins \cite{bib:Tillmann2015,bib:Bozinovic2013, bib:Nicolas2014, bib:Humphreys2013, bib:Donohue2013}, are present. When multiple photons are present, they experience quantum interference when all quantum labels are the same.

@@

\begin{align}
b_i^{\dag}=\sum_{j=1}^d U_{ij}a_j^\dag \ ,
\end{align}
where $a_j^\dag$ and $b_i^\dag$ are the creation operators of a single input and output photon in the $j$-th and $i$-th modes respectively, and $U\in SU(d)$. All such transformation can be expressed as sequence of beamsplitters and waveplates \cite{bib:Reck1994}. In the case when photons have additional labels, for instance, if they have internal labels on top of spatial labels, it is also possible to derive an analogous decomposition, known as a cosine-sine decomposition \cite{bib:Dhand2015}, that realizes the unitary transformation on the photons into a sequence of beamsplitters and internal transformations. Toolkits using group theory are being developed to deal with partial distinguishabilities among interfering photons \cite{bib:Tan2013,bib:deGuise2014,bib:deGuise2015}. Others use
quantum-to-classical transitions to explain multiparticle interference \cite{bib:Ra2013}. \sihui{Experimental implementation [Walmsley, Jennewein]}

@@

\textbf{Need to edit the above together with what follows...}

It was first conceived for a single-photon passing through a Mach-Zedner interferometer with its two paths having different lengths \cite{bib:Brendel99}. If the photon passes through the shorter (longer) path, then it will arrive at the output port in an ``early" (``late") time bin labelled $\ket{e}$ ($\ket{\ell}$). Thus, photon will exit the interferometer in a state that is a superposition of these two states. Owing to difficulties in implementing qubit operations in this basis, it was at that time mostly used for demonstrating quantum communication over long distance \cite{bib:Thew02,bib:Marcikic04}. With the advent of faster optical components, and single-photon detectors, it has become feasible to perform any single qubit operations, and a post-selected CPHASE gate on time-bin qubits \cite{bib:Humphreys2013}. These gates form an universal set for quantum computing. At the same time, an ultrafast measurement technique for recovering time-bin qubits was also demonstrated \cite{bib:Donohue2013}.

In Humphrey {\it et al.} \cite{bib:Humphreys2013}, the time-bin qubits have a polarization register which is used to toggle between the mode in which qubits are stored and transmitted, {\it i.e.} the ``register" mode, and the mode in which qubits are manipulated, {\it i.e.} the ``processing" mode using a polarization rotation. Other operations needed are a displacement operation moves a time-bin qubit in the processing polarization in time relative to other qubits, a partial polarization rotation between the register and processing polarizations to couple them, and a photon number measurement in a bin. The polarization rotation and polarization coupling operations were implemented using a fast integrated optical switch based on polarization-sensitive cross-phase modulation with a switching time of less than 10 ps. The displacement was implemented using a calcite crystal of a size that is of a path-length difference equal to an integer multiple of the time-bin separation. Using these operations and post-selection using a single-photon detectors, the authors were able to perform a heralded CPHASE gate \cite{bib:KLM01}.

In Donohue {\it et al.} \cite{bib:Donohue2013}, time bins were converted into frequency bins using a nonlinear optical process known as sum-frequency generation (SFG) by pumping them on a crystal with a laser beam. The waveforms of the input photons and the pump beam are chirped in opposite directions in frequency, so that the output is a set of two peaks separated in frequency by an amount proportional to the time separation of the time bins. The output can then be measured with detectors that are slow compared to this separation. This is unlike conventional time-bin measurements which involves sending a time-bin state through an unbalanced Mach-Zedner interferometer matched to a bin separation to produce three output pulses that are separated by the time delay, and then having to resolve the middle pulse in time.

Apart from these promising advances in produce time-bin qubit gates and ultrafast time-bin measurements, scalable networks using time-bin encoding are also possible using a loop-based architecture \cite{bib:Motes14}. This is discussed in more detail in Section \ref{sec:timebins}.

@@

The task of implementing an arbitrary quantum computation on linear optics comes down to implementing an arbitrary $n\times n$ unitary matrix. If a non-unitary transformation is desired, it can be embedded within a unitary matrix with larger dimensions. An algorithm for expressing an arbitrary unitary matrix {\it exactly} in terms of a sequence of $\mathcal{O}(n^2)$ beamsplitters and phase-shifters exists \cite{bib:Reck1994}. Alternatively, Mach-Zedner interferometer can also be used as building blocks instead of beamsplitters and phase shifters \cite{bib:Reck1994, bib:Englert2001}. Later, it has been shown that any nontrivial beam splitter, that does more than swapping modes around or add phases to them, is universal for linear optics \cite{bib:Bouland2014}. However, they do not provide a construction for arbitrary unitaries.

If the linear optical transformations can be realized on various degrees of freedom of light, then it is possible to realize a $n\times n$ arbitrary unitary transformation, where $n=n_s n_p$ for $n_s$ spatial modes, and $n_p$ internal modes, by a sequence of $\mathcal{O}(n_s^2 n_p)$ beamsplitters and $\mathcal{O}(n_s^2)$ internal transformations \cite{bib:Dhand2015}. Their approach reduces the required number of beamsplitters but increases the total number of optical elements needed increases by a factor of 2.

@@

In many practical situations, the structure of a linear optical device in terms of its constituent beamsplitters and phase shifters is known once it is built. However, owing to manufacturing imperfections, a precise characterization of these devices is still needed post-production. One of the ways, this can be done is via a quantum process tomography \cite{bib:Mitchell03,bib:Obrien04,bib:Lobino08,bib:Saleh11} . However, quantum process tomography is an expensive method in terms of number of measurements required to characterize the network, and it becomes impractical for large optical networks which can now be as large as 900 modes (check citations for number of modes). Alternative characterization protocols have been developed using quantum interference of various quantum light sources \cite{bib:Laing12,bib:Rahimi-Keshari13} in the linear optical device.

Generally, the unitary matrix of the $d\times d$ linear optical device are complex numbers $U_{ij}=r_{ij}e^{i\theta_{ij}}$, where $0 \leq r_{ij}\leq 1$, and $0\leq \theta_{ij}\leq 2\pi$. The scheme in \cite{bib:Laing12} relies on injecting one-photon and two-photon states into the linear optical network with correlated photon detection. First, they note some equivalencies: two unitaries $U$ and $U'$ are equivalent if there exist two diagonal unitary matrices $D_1^U$ and $D_2^U$ such that $U'=D^U_1 U D^U_2$, because these diagonal matrices are regarded as unknown and trivial phases on the input and output ports of the network, to which the one-photon and two-photon data are insensitive to. This reduces the first row and first column elements to real numbers, i.e. $\theta_{1j}=\theta_{i1}=0$. Second, the photon statistics remain unchanged under the complex conjugation of $U$. Thus, the imaginary part of $M_{2,2}$ must be non-negative. Then assuming that the first two rows and columns are non-vanishing, and that there is no total loss in the interferometer, the matrix to be recovered is
\begin{align}
U=\left (\begin{array}{cccc}
r_{11} & r_{12} & \ldots & r_{1m} \\
r_{21} & r_{22}e^{i\theta_{22}} &\ldots & r_{2m}e^{i\theta_{2m}}\\
\vdots & \vdots & \ddots & \vdots \\
r_{m1} & r_{m2}e^{i\theta_{m2}} & \ldots & r_{mm}e^{i\theta_{mm}}
 \end{array}\right ) \ .
\end{align}
In the paper, they showed that it is possible to write the paramaters $r_{ij}$, and $\theta_{ij}$ for $i,j\geq 2$ in terms of the $2m-1$ real parameters of the first column and row, and the visibility of two-photon inputs. The remaining $2m-1$ real parameters can be found via one-photon 
transmissions.  An increased accuracy of the characterization is possible by estimating and correcting systematic errors that arise due to mode mismatch \cite{Dhand16}. Others have used numerical methods to find the closest parameters that yield the observed visibilities \cite{bib:Spagnolo16,bib:Tillmann16}.

Another characterization method was presented that is similar to \cite{bib:Laing12} with the important exception that coherent states are to be used instead \cite{bib:Rahimi-Keshari13,bib:Heilmann15}. Such states are produced by a standard laser source, thus reducing the resource needed. The $r_{jk}$ terms are found by the square root of the ratios of output intensity at the $k$th port to the input intensity at the $j$th port. The remaining phases $\theta_{ij}$ are found by the interference pattern given by a two-mode coherent state $\ket{\alpha_1}\ket{\alpha_2}$ created by splitting a single coherent state on a 50-50 beamsplitter, and then imparting a relative phase, $\phi$, between them. The states $\ket{\alpha_1}$ and $\ket{\alpha_2}$ are input into port 1, and $j$ respectively. The output intensity at the $j$th port is 
\begin{align}
I_k=I(r_{1k}^2+r_{jk}^2+2 r_{1k}r_{kj}\cos(\phi+\theta_{jk})) \ ,
\end{align}
where $\theta_{jk}=0$ for $k=1$. By scanning the phase shift $\phi$ and then locating the maximum value of $I_k$ for $j=2,\ldots, m$, all unknown phases can be found via $\theta_{jk}=2\pi-\phi$. An elegant extension of the scheme of Rahimi-Keshari {\it et al.} removes the need for precise control of the phase shift $\phi$ \cite{bib:Heilmann15} by suggesting instead to plot the output intensity $I_k$ with respect to the input intensity $I$. In time, the natural drift in the laser source will cause this plot to trace out an ellipse, known as a Lissajous figure. whose orientation and direction of evolution will give the phase $\theta_{jk}$ and its sign respectively.

@@

@@

@@

Sources of single photons for applications of quantum information processing are 
separated into two main categories \cite{bib:Kok05}; those produced by spontaneous 
parametric down-conversion (SPDC), and those by solid-state emitters in a cavity. 
Both categories have seen some leaps recently in their record production of numbers 
of single photons while preserving good qualities in their single photons. To 
date, SPDC has been able to produce up to ten entangled photons 
\cite{bib:WangChen16,bib:Chen17}, and five-photon quantum interference has been 
reported using a quantum dot emitter in a microcavity \cite{bib:WangHe16}.

In SPDC, a nonlinear crystal with a large $\chi^{(2)}$ is pumped with a strong 
laser and with a small probability, the pump beam is absorbed by the crystal to 
produce two beams of lower energy known as the signal and idler beams. Owing to 
conservation of energy and momentum, the two beams have spatio-temporal 
correlations that can be engineered to produce twin-beam states which have 
perfectly correlated photon numbers. If a single photon were to be detected in one 
of the beams, and then it is certain that the other beam would be in the state of a 
single photon. Some commonly used nonlinear crystals for SPDC are beta barium 
borate (BBO), periodically poled lithium niobate (PPLN), and periodically poled 
potassium titanyl phosphate (PPKTP). Recently, techniques using bismuth triborate 
(BiBO) have improved to the extent of becoming one of the record-holders of single-
photon production \cite{bib:WangChen16}.

Solid-state single photon sources come from semiconducting nanostructures and 
nitrogen valencies (NV) centers in diamond. Both types are versatile and efficient 
sources of single photons, however, the photons they produce have suffered from 
lack of indistinguishability that is necessary for quantum information processing.
Recent developments in the former show promising developments. Notably, much 
progress has been made using resonant excitation of quantum dots. Laser pulses are 
used to excite their electronic resonance and trigger the emission of single 
photons \cite{bib:Muller07,bib:Vamivakas09, bib:Flagg09,bib:Ates09, bib:Dirk10,
bib:1748,bib:Jayakumar13,bib:Wei14,bib:Muller14,bib:Unsleber15,
bib:237403,bib:Sweeney14}. An architecture that embeds the quantum dot in a 
micropillar cavity with the same resonant frequency as the dot exploits the 
Purcell effect to achieve higher single-photon production efficiency 
†\cite{bib:Ding16,bib:213601}. 
Such resonant excitation of quantum dots overcome the homogenous broadening of the 
excited state that causes degradation of photon purity and hence improves
indistinguishability. Using a time-bin loop-based architecture \cite{bib:Motes14} 
on a stream of single photons produced by such a quantum dot, the same group was 
able to demonstrate quantum interference between five indistinguishable photons 
\cite{bib:WangChen16}.

@@

\sihui{ Discuss producing Bell pairs using SPDC, and transforming them into GHZ and cluster states.}

An EPR pair is a pair of qubits that are in a Bell state together, that is, one of the following states: $\{\ket{0}_A\ket{0}_B\pm\ket{1}_A\ket{1}_B,\{\ket{1}_A\ket{0}_B\pm\ket{1}_A\ket{0}_B\}$. Such a pair of qubits forms the simplest example of entangled states, and its production from spontaneous parametric down-conversion (SPDC) have been the mainstay of entanglement source for quantum optical processing \cite{bib:Kim01,bib:Kim03,bib:Brida07,bib:Dayan07}. In some applications, it may be desired to have the EPR pairs conditionally prepared, {\it i.e.} successfully prepared only under certain measurement outcomes of auxiliary modes. Some theoretical proposals have been proposed \cite{bib:Pittman03,bib:Sliwa03,bib:Walther07}, and in some cases demonstrated \cite{bib:Wagenknecht10,bib:Barz10}.

@@

@@

@@

@@

@@

@@

\cite{bib:Motes14}

@@

The workings of the scheme can be thought of as follows: the inner loop allows arbitrary beamsplitter operations between neighbouring time-bins; the outer loop does nothing interferometric, but rather enables the pulse-train to undergo as many applications of the inner loops as necessary. It then follows that this scheme is universal for linear optics, as a sufficient number of beamsplitter operations between neighbouring modes enables universal decompositions, such as that by \cite{Reck}.

As a simple example of how such time-bin encoding maps to other encodings, in Fig.~\ref{fig:PBS_TB} we show the isomorphism between the polarising beamsplitter operation and a pairwise temporal beamsplitter operation.

@@

\subsection{Measurement}

\subsubsection{Photodetection}

\comment{Discuss number-resolved and bucket detectors, multiplexed detection, APDs, current micropillar detectors}

\subsubsection{Homodyning}

@@

\subsection{Encrypted quantum computation}

\section{State of the art}

\comment{Discuss where experiments are at at the moment}

@@
